The state learner -- a super learner for right-censored data

### 2024-05-27 tex/journal-version/manuscript.tex
renaming
reformat manuscript

%%
%% Copyright 2022 OXFORD UNIVERSITY PRESS
%%
%% This file is part of the 'oup-authoring-template Bundle'.
%% ---------------------------------------------
%%
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%%
%% The list of all files belonging to the 'oup-authoring-template Bundle' is
%% given in the file `manifest.txt'.
%%
%% Template article for OXFORD UNIVERSITY PRESS's document class `oup-authoring-template'
%% with bibliographic references
%%

%%%CONTEMPORARY%%%
% \documentclass[unnumsec,webpdf,contemporary,large]{oup-authoring-template}%
\documentclass[numsec,webpdf,contemporary,medium,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
%\documentclass[unnumsec,webpdf,contemporary,medium]{oup-authoring-template}
%\documentclass[unnumsec,webpdf,contemporary,small]{oup-authoring-template}

%%%MODERN%%%
%\documentclass[unnumsec,webpdf,modern,large]{oup-authoring-template}
% \documentclass[unnumsec,webpdf,modern,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
%\documentclass[unnumsec,webpdf,modern,medium]{oup-authoring-template}
%\documentclass[unnumsec,webpdf,modern,small]{oup-authoring-template}

%%%TRADITIONAL%%%
%\documentclass[unnumsec,webpdf,traditional,large]{oup-authoring-template}
% \documentclass[unnumsec,webpdf,traditional,large,namedate]{oup-authoring-template}% uncomment this line for author year citations and comment the above
%\documentclass[unnumsec,namedate,webpdf,traditional,medium]{oup-authoring-template}
%\documentclass[namedate,webpdf,traditional,small]{oup-authoring-template}

\onecolumn % for one column layouts

%\usepackage{showframe}

% \graphicspath{{Fig/}}

% line numbers
%\usepackage[mathlines, switch]{lineno}
%\usepackage[right]{lineno}

@@

%% OLD

% \usepackage{natbib, dsfont,
% amssymb,soul,xcolor,amsmath,graphicx,verbatim,pgfplots,tikz,prodint,booktabs}

\usepackage{dsfont, booktabs}

@@

%% change enumerate labels
\renewcommand{\labelenumi}{\normalfont (\alph{enumi})}

@@

% \bibliographystyle{rss}

@@

\newcommand{\includeFigCond}[2][]{
  \ifx\nofig\undefined %
    \includegraphics[#1]{#2} %
  \else %
    \texttt{#2} %
  \fi %
}

@@

@@

%% NEW

\journaltitle{Journal of the Royal Statistical Society Series B: Statistical
  Methodology} %
\DOI{DOI HERE} %
\copyrightyear{xxxx} %
\pubyear{yyyy} \access{Advance Access Publication Date: Day Month Year}
\appnotes{Paper}

\firstpage{1}

%\subtitle{Subject Section}

\title[The state learner]{The state learner -- a super learner for right-censored data}

\author[1,$\ast$]{Anders Munch}
\author[1]{Thomas A.\ Gerds}

\authormark{Munch et al.}

\address[1]{\orgdiv{Section of Biostatistics}, \orgname{University of
    Copenhagen}, \orgaddress{\street{Østerfarigmagsgade 5}, \postcode{2100}, \state{Copenhagen}, \country{Denmark}}}

\corresp[$\ast$]{Address for correspondence: \href{email:email-id.com}{a.munch@sund.ku.dk}}

@@

\abstract{In survival analysis, prediction models are needed as
  stand-alone tools and in applications of causal inference to
  estimate nuisance parameters. The super learner is a machine
  learning algorithm which combines a library of prediction models
  into a meta learner based on cross-validated loss. In right censored
  data, the choice of the loss function and the estimation of the
  expected loss need careful consideration. We introduce the state
  learner, a new super learner for survival analysis, which
  simultaneously evaluates libraries of prediction models for the
  event of interest and the censoring distribution. The state learner
  can be applied to all types of survival models, works in the
  presence of competing risks, and does not require a single
  pre-specified estimator of the conditional censoring
  distribution. We establish an oracle inequality for the state
  learner and investigate its performance through numerical
  experiments. We illustrate the application of the state learner with
  prostate cancer data, as a stand-alone prediction tool, and, for
  causal inference, as a way to estimate the nuisance parameter models of
  a smooth statistical functional.
  % We illustrate how the state learner allows us to
  % estimate causal effects in a competing risks setting without having
  % to pre-specify models for neither the cause-specific hazards of
  % interest nor the censoring distribution.
}

\keywords{Competing risks, cross-validation, loss based estimation,
  right-censored data, super learner}

@@

@@

We introduce our notation and framework in
Section~\ref{sec:framework}. In Section~\ref{sec:super-learning} we
define general super learning for right-censored data.
% , and in
% Section~\ref{sec:relat-liter-exist} we review existing methods.
Section~\ref{sec:super-learner-simple} introduces the state learner,
and Section~\ref{sec:theor-results-prop} provides theoretical
guarantees. In Section~\ref{sec:targeted-learning} we discuss the use
of the state learner in the context of targeted learning. We report
results of our numerical experiments in Section~\ref{sec:numer-exper}
and analyze a prostate cancer data set in
Section~\ref{sec:real-data-appl}.  Section~\ref{sec:discussion}
contains a discussion of the merits and limitations of our
proposal. Appendices~\ref{sec:proof-proposition}
and~\ref{sec:state-learner-with} contain proofs.

@@

The function \( F \) describes the conditional state occupation
probabilities of the multi-state process \(\eta\). We construct a
super learner for \( F \). The target of this super learner is the
function-valued parameter $\Psi(P) = F$ which is identified through
equation~(\ref{eq:F-def}).
Under conditional independent censoring each quadruple
$(\Lambda_{1}, \Lambda_{2}, \Gamma, H)$ characterises a distribution
\(P\in\mathcal P\), c.f.\ equation~\eqref{eq:parametrizeP}, which in turn
determines \( (F, H) \).
Hence, a learner for \( F \) can be constructed from learners for
\( \Lambda_1 \), \( \Lambda_2 \), and $\Gamma$ as follows:

@@

  & = P(\tilde{T} \leq t, \tilde{D}=1 \mid X=x)
    = \int_0^t \exp{{\{-\Lambda_{1}(s \mid x)-\Lambda_{2}(s \mid x) - \Gamma(s \mid x)\} }}  \Lambda_{1}(\diff s \mid x),
  & = P(\tilde{T} \leq t, \tilde{D}=2 \mid X=x)
    = \int_0^t \exp{\{-\Lambda_{1}(s \mid x)-\Lambda_{2}(s \mid x) - \Gamma(s \mid x)\} }  \Lambda_{2}(\diff s \mid x),
    P(\tilde{T} \leq t, \tilde{D}=0 \mid X=x)
    = \int_0^t \exp{\{-\Lambda_{1}(s \mid x)-\Lambda_{2}(s \mid x) - \Gamma(s \mid x)\} }  \Gamma(\diff s \mid x),

@@

    = 1- F(t, 1, x) - F(t, 2, x)- F(t, -1, x).

@@

@@

    \intertext{where in correspondence with  the relations in equation \eqref{eq:transition},} 
    \phi_{a_1,a_2, b}(\data_n)(t,1,x) &= \int_0^t \exp{\{-a_1(\data_n)(s \mid x)-a_2(\data_n)(s \mid x) - b(\data_n)(s \mid x)\} }  a_1(\data_n)(\diff s \mid x),\\
  \phi_{a_1,a_2, b}(\data_n)(t,2,x) &= \int_0^t \exp{\{-a_1(\data_n)(s \mid x)-a_2(\data_n)(s \mid x) - b(\data_n)(s \mid x)\} }  a_2(\data_n)(\diff s \mid x),\\
  \phi_{a_1,a_2, b}(\data_n)(t,-1,x) &= \int_0^t \exp{\{-a_1(\data_n)(s \mid x)-a_2(\data_n)(s \mid x) - b(\data_n)(s \mid x)\} }  b(\data_n)(\diff s \mid x),\\
  \phi_{a_1,a_2, b}(\data_n)(t,0,x) &= 1
                                      - \phi_{a_1,a_2, b}(\data_n)(t,1,x)
                                      - \phi_{a_1,a_2, b}(\data_n)(t,2,x)
    -\phi_{a_1,a_2, b}(\data_n)(t,-1,x)

@@

@@

\begin{corollary}
  \label{cor:oracle-prop}
  For all \(P_0\in\mathcal{P}\), \( n \in \N \), \( k \in \{1, \dots, K\} \),
  and $\delta>0$,
  \begin{align*}
    \E_{P_0}{\left[ \Vert \hat{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]}
    & \leq (1 + 2\delta)
      \E_{P_0}{\left[ \Vert \tilde{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]}
    % \\
    % & \quad
      + (1+ \delta) 16   K \tau
      \left(
      13 + \frac{12}{\delta}
      \right)
      \frac{\log(1 + |\mathcal{F}_n|)}{n}.
  \end{align*}
\end{corollary}
\begin{proof}
  See Appendix~\ref{sec:proof-proposition}.
\end{proof}

@@

\begin{corollary}
  \label{cor:asymp-cons}
  Assume that \( |\mathcal{F}_n| = \bigO(n^q)\), for some \( q \in \N \)
  and that there exists a sequence \( \phi_n \in \mathcal{F}_n \),
  \( n \in \N \), such that
  \( \E_{P_0}{\left[ \Vert \phi_n(\data_n^{-k}) - F_{0} \Vert_{P_0}^2 \right]} =
  \bigO(n^{-\alpha}) \), for some \( \alpha\leq 1 \).
  \begin{enumerate}[1.]
  \item If $\alpha=1$ then
    \( \E_{P_0}{\left[ \Vert \hat{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2
      \right]} = \bigO(\log(n)n^{-1}) \).
  \item If $\alpha<1$ then
    \( \E_{P_0}{\left[ \Vert \hat{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]} =
    \bigO(n^{-\alpha}) \).
  \end{enumerate}
\end{corollary}
\begin{proof}
  See Appendix~\ref{sec:proof-proposition}.
\end{proof}

@@

Using equation~(\ref{eq:surv-def}) we may write
\( \theta_{{\tau}}(Q) = \tilde{\Psi}_{\tau}^0(\Lambda_1, \Lambda_2, \mu) \), where
\begin{equation}
  \label{eq:1}    
  \begin{split}
    \tilde{\Psi}_{\tau}^0(\Lambda_1, \Lambda_2, \mu)
    & =
      \int_{\mathcal{W}} 
      \int_0^{\tau}
      \exp{\{{-\Lambda_1(s- \mid w, 1)-\Lambda_2(s- \mid w, 1)}\}}  \Lambda_1(\diff s \mid
      w, 1)
      \mu(\diff w)
    \\
    &  \quad
      -\int_{\mathcal{W}} 
      \int_0^{\tau}
      \exp{\{{-\Lambda_1(s- \mid w, 0)-\Lambda_2(s- \mid w, 0)}\}}  \Lambda_1(\diff s \mid w, 0)
      \mu(\diff w).
  \end{split}
\end{equation}
The efficient influence function for the parameter $\tilde{\Psi}_{\tau}$ depends
on the set \( (\Lambda_1, \Lambda_2, \Gamma, \pi) \) of nuisance parameters.
We define
\begin{align*}
  \omega_a(A,W; \pi)
  &=  \frac{% (-1)^{a+1}
    \1{\{A=a\}}}{\pi(W)^{a}(1-\pi(W))^{1-a}},
  \\
  g(t, A, W; \Lambda_1, \Lambda_2)
  & = \int_0^{t}
    \exp{\{ {-\Lambda_1(s- \mid W, A)-\Lambda_2(s- \mid W, A)} \}}  \Lambda_1(\diff s \mid
    W, A),
  \\  
  M_j(\diff t \mid A, W;  \Lambda_j  )
  & = N_j(\diff t) -
    \1{\{\tilde{T} \geq t\}} \Lambda_j(\diff t \mid W, A),
    \quad j \in \{1,2\},
    \intertext{and}
    M(\diff t \mid A, W;  \Lambda_1, \Lambda_2  )
  & = M_1(\diff t \mid A, W;  \Lambda_1  ) +
    M_2(\diff t \mid A, W;  \Lambda_2  ).
\end{align*}
The efficient influence function can now be written as
\citep{van2003unified,jewell2007non,rytgaard2022targeted},
\begin{equation*}
  \psi_{\tau}(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
  = \psi_{\tau}^1(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
  - \psi_{\tau}^0(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
  -\tilde{\Psi}_{\tau}^0(\Lambda_1, \Lambda_2, \mu),
\end{equation*}
where
\begin{equation}
  \label{eq:5}
  \begin{split}
    & \psi_{\tau}^a(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
    \\
    & =
      \omega_a(A,W; \pi)
      \int_0^{\tau} \exp{\{{\Gamma(t- \mid A, W)}\}}   
      M_1(\diff t \mid A, W; \Lambda_1)
    \\
    & \quad
      -
      \omega_a(A,W; \pi)
      g(\tau, A, W; \Lambda_1, \Lambda_2)
      \int_0^{\tau}
      \exp{\{{[\Gamma+\Lambda_1 + \Lambda_2](t- \mid A, W)}\}}
      M(\diff t \mid A, W; \Lambda_1, \Lambda_2)
    \\
    & \quad
      +
      \omega_a(A,W; \pi)      
      \int_0^{\tau}
      g(t, A, W; \Lambda_1, \Lambda_2)
      \exp{\{{[\Gamma+\Lambda_1 + \Lambda_2](t- \mid A, W)}\}}
      M(\diff t \mid A, W; \Lambda_1, \Lambda_2)
    \\
    & \quad + g(\tau, a, W; \Lambda_1, \Lambda_2).
  \end{split}
\end{equation}
Equations~(\ref{eq:1}) and~(\ref{eq:5}) allow us to construct a one-step
estimator by using the definition given in equation~(\ref{eq:one-step-def}),
which gives the estimator
\begin{equation}
  \label{eq:one-step-comp-ate}
  \begin{split}
    \hat{\Psi}_{\tau,\text{OS}}(\data_n)
    = &
        \tilde{\Psi}_{\tau}^0(\hat{a}_1(\data_n), \hat{a}_2(\data_n),
        \hat{\mu}_n(\data_n))
    \\
      &
        +
        \empmeas{[\psi_{\tau}(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
        \hat{b}_n(\data_n), \hat{\pi}_n(\data_n))]}
    \\
    = &
        \empmeas{[\psi_{\tau}^1(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
        \hat{b}_n(\data_n), \hat{\pi}_n(\data_n))]}
    \\
      &
        - \empmeas{[\psi_{\tau}^0(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
        \hat{b}_n(\data_n), \hat{\pi}_n(\data_n))]}.
  \end{split}
\end{equation}

@@

\begin{appendices}

@@

\end{appendices}

\bibliographystyle{abbrvnat}
\bibliography{bib.bib}

>>>>

\documentclass[12pt]{article}
\usepackage[a4paper,left=2.8cm, right=2.8cm, top=3cm, bottom=3cm]{geometry}

%% Packages
\usepackage{amsthm, dsfont, amssymb,soul,xcolor,
  % amsfonts,
  enumitem,
  amsmath,graphicx,verbatim,pgfplots,tikz,prodint,booktabs}

\usepackage[authoryear]{natbib}
\bibliographystyle{abbrvnat}

% % Handling new lines
% \setlength{\parskip}{1em}
% \setlength{\parindent}{0em}

% % Handling space after sections
% \usepackage{titlesec}
% \titlespacing*{\section}{0em}{2em}{0em}
% \titlespacing*{\subsection}{0em}{2em}{0em}
% \titlespacing*{\subsubsection}{0em}{2em}{0em}

% % No spacing after in start of list
% \setlist[itemize]{topsep=0pt}
% \setlist[enumerate]{topsep=0pt}

\definecolor{linkcolor}{rgb}{0, 0, 0.54}
\usepackage[colorlinks,allcolors=linkcolor]{hyperref}

@@

@@

%% change enumerate labels
% \renewcommand{\labelenumi}{\normalfont (\alph{enumi})}

%% Font size etc
\linespread{1.65}

\usepackage{lineno}
\linenumbers

@@

@@

\RequirePackage{graphicx}
% \RequirePackage{caption}
% \RequirePackage{array}
% \RequirePackage{color}
% \RequirePackage{xcolor}
% \RequirePackage{amssymb}
% \RequirePackage{flushend}
% \RequirePackage{stfloats}
% \RequirePackage[figuresright]{rotating}
% \RequirePackage{chngpage}
% \RequirePackage{totcount}
% \RequirePackage{fix-cm}
\newcommand{\includeFigCond}[2][]{
  \ifx\nofig\undefined %
    \includegraphics[#1]{#2} %
  \else %
    \texttt{#2} %
  \fi %
}

@@

%% Title
\title{The state learner -- a super learner for right-censored data}

\usepackage{authblk}
\author[1]{Anders Munch}
\author[1]{Thomas~A.~Gerds}
\affil[1]{Section of Biostatistics, University of Copenhagen}

@@

% \authormark{Munch et al.}

% \address[1]{\orgdiv{Section of Biostatistics}, \orgname{University of
%     Copenhagen}, \orgaddress{\street{Østerfarigmagsgade 5}, \postcode{2100}, \state{Copenhagen}, \country{Denmark}}}

% \corresp[$\ast$]{Address for correspondence: \href{email:email-id.com}{a.munch@sund.ku.dk}}

@@

% \keywords{Competing risks, cross-validation, loss based estimation,
%   right-censored data, super learner}

@@

\begin{abstract}
  In survival analysis, prediction models are needed as stand-alone tools and in
  applications of causal inference to estimate nuisance parameters. The super
  learner is a machine learning algorithm which combines a library of prediction
  models into a meta learner based on cross-validated loss. In right censored
  data, the choice of the loss function and the estimation of the expected loss
  need careful consideration. We introduce the state learner, a new super
  learner for survival analysis, which simultaneously evaluates libraries of
  prediction models for the event of interest and the censoring distribution.
  The state learner can be applied to all types of survival models, works in the
  presence of competing risks, and does not require a single pre-specified
  estimator of the conditional censoring distribution. We establish an oracle
  inequality for the state learner and investigate its performance through
  numerical experiments. We illustrate the application of the state learner with
  prostate cancer data, as a stand-alone prediction tool, and, for causal
  inference, as a way to estimate the nuisance parameter models of a smooth
  statistical functional.
\end{abstract}

@@

We introduce our notation and framework in
Section~\ref{sec:framework}. In Section~\ref{sec:super-learning} we
define general super learning for right-censored data.
% , and in
% Section~\ref{sec:relat-liter-exist} we review existing methods.
Section~\ref{sec:super-learner-simple} introduces the state learner,
and Section~\ref{sec:theor-results-prop} provides theoretical
guarantees. In Section~\ref{sec:targeted-learning} we discuss the use
of the state learner in the context of targeted learning. We report
results of our numerical experiments in Section~\ref{sec:numer-exper}
and analyze a prostate cancer data set in
Section~\ref{sec:real-data-appl}.  Section~\ref{sec:discussion}
contains a discussion of the merits and limitations of our
proposal. Appendices~\ref{sec:proof-proposition}
and~\ref{sec:state-learner-with} contain proofs.
An implementation of the state learner is available at
\url{https://github.com/amnudn/statelearner} along with codes for reproducing
our numerical experiments.

@@

The function \( F \) describes the conditional state occupation probabilities of
the multi-state process \(\eta\). We construct a super learner for \( F \). The
target of this super learner is the function-valued parameter $\Psi(P) = F$
which is identified through equation~(\ref{eq:F-def}). Under conditional
independent censoring each quadruple $(\Lambda_{1}, \Lambda_{2}, \Gamma, H)$
characterises a distribution \(P\in\mathcal P\), c.f.\
equation~\eqref{eq:parametrizeP}, which in turn determines \( (F, H) \). Hence,
a learner for \( F \) can be constructed from learners for \( \Lambda_1 \),
\( \Lambda_2 \), and $\Gamma$ as follows:

@@

@@

      =
      \exp{{\{-\Lambda_{1}(t \mid x)-\Lambda_{2}(t \mid x) - \Gamma(t \mid x)\}
      }},
      = P(\tilde{T} \leq t, \tilde{D}=1 \mid X=x)
      = \int_0^t F(s-, 0, x)  \Lambda_{1}(\diff s \mid x),
      = P(\tilde{T} \leq t, \tilde{D}=2 \mid X=x)
      = \int_0^t  F(s-, 0, x)  \Lambda_{2}(\diff s \mid x),
      = P(\tilde{T} \leq t, \tilde{D}=0 \mid X=x)
      = \int_0^t F(s-, 0, x)  \Gamma(\diff s \mid x).

@@

% \begin{equation}\label{eq:transition}
%   \begin{split}
%   F(t, 1, x)
%     &
%       % = P(\tilde{T} \leq t, \tilde{D}=1 \mid X=x)
%     = \int_0^t \exp{{\{-\Lambda_{1}(s \mid x)-\Lambda_{2}(s \mid x) - \Gamma(s \mid x)\} }}  \Lambda_{1}(\diff s \mid x),
%   \\
%   F(t, 2, x)
%     &
%       % = P(\tilde{T} \leq t, \tilde{D}=2 \mid X=x)
%       = \int_0^t \exp{\{-\Lambda_{1}(s \mid x)-\Lambda_{2}(s \mid x) - \Gamma(s \mid x)\} }  \Lambda_{2}(\diff s \mid x),
%     \\
%     F(t, -1, x)
%     &
%       % = P(\tilde{T} \leq t, \tilde{D}=0 \mid X=x)
%       = \int_0^t \exp{\{-\Lambda_{1}(s \mid x)-\Lambda_{2}(s \mid x) - \Gamma(s \mid x)\} }  \Gamma(\diff s \mid x),
%     \\
%     F(t, 0, x)
%     &
%       = P(\tilde{T} > t \mid X= x)
%       = 1- F(t, 1, x) - F(t, 2, x)- F(t, -1, x).
%   \end{split}
% \end{equation}

@@

    \intertext{where in correspondence with  the relations in equation
    \eqref{eq:transition},}
    \phi_{a_1,a_2, b}(\data_n)(t,0,x)
  &= \exp{\{-a_1(\data_n)(s \mid x)-a_2(\data_n)(s \mid x) - b(\data_n)(s \mid
    x)\} },
  \phi_{a_1,a_2, b}(\data_n)(t,1,x)
  &= \int_0^t
    \phi_{a_1,a_2, b}(\data_n)(s-,0,x)  a_1(\data_n)(\diff s \mid x),
  \phi_{a_1,a_2, b}(\data_n)(t,2,x)
  &= \int_0^t\phi_{a_1,a_2, b}(\data_n)(s-,0,x)  a_2(\data_n)(\diff s \mid x),
  \phi_{a_1,a_2, b}(\data_n)(t,-1,x)
  &= \int_0^t \phi_{a_1,a_2, b}(\data_n)(s-,0,x)  b(\data_n)(\diff s \mid x).

@@

% \begin{align*}
%   \mathcal{F}(\mathcal{A}_1, \mathcal{A}_2, \mathcal{B})
%   &= \{ \phi_{a_1,a_2, b} : a_1 \in \mathcal{A}_1, a_2 \in \mathcal{A}_2, b \in \mathcal{B}\},
%     \intertext{where in correspondence with  the relations in equation \eqref{eq:transition},} 
%     \phi_{a_1,a_2, b}(\data_n)(t,1,x) &= \int_0^t \exp{\{-a_1(\data_n)(s \mid x)-a_2(\data_n)(s \mid x) - b(\data_n)(s \mid x)\} }  a_1(\data_n)(\diff s \mid x),\\
%   \phi_{a_1,a_2, b}(\data_n)(t,2,x) &= \int_0^t \exp{\{-a_1(\data_n)(s \mid x)-a_2(\data_n)(s \mid x) - b(\data_n)(s \mid x)\} }  a_2(\data_n)(\diff s \mid x),\\
%   \phi_{a_1,a_2, b}(\data_n)(t,-1,x) &= \int_0^t \exp{\{-a_1(\data_n)(s \mid x)-a_2(\data_n)(s \mid x) - b(\data_n)(s \mid x)\} }  b(\data_n)(\diff s \mid x),\\
%   \phi_{a_1,a_2, b}(\data_n)(t,0,x) &= 1
%                                       - \phi_{a_1,a_2, b}(\data_n)(t,1,x)
%                                       - \phi_{a_1,a_2, b}(\data_n)(t,2,x)
%     -\phi_{a_1,a_2, b}(\data_n)(t,-1,x)
% \end{align*}

@@

\begin{corollary}
  \label{cor:oracle-prop}
  For all \(P_0\in\mathcal{P}\), \( n \in \N \), \( k \in \{1, \dots, K\} \),
  and $\delta>0$,
  \begin{align*}
    \E_{P_0}{\left[ \Vert \hat{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]}
    & \leq (1 + 2\delta)
      \E_{P_0}{\left[ \Vert \tilde{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]}
    \\
    & \quad
      + (1+ \delta) 16   K \tau
      \left(
      13 + \frac{12}{\delta}
      \right)
      \frac{\log(1 + |\mathcal{F}_n|)}{n}.
  \end{align*}
\end{corollary}
\begin{proof}
  See Appendix~\ref{sec:proof-proposition}.
\end{proof}

@@

\begin{corollary}
  \label{cor:asymp-cons}
  Assume that \( |\mathcal{F}_n| = \bigO(n^q)\), for some \( q \in \N \) and
  that there exists a sequence \( \phi_n \in \mathcal{F}_n \), \( n \in \N \),
  such that
  \( \E_{P_0}{\left[ \Vert \phi_n(\data_n^{-k}) - F_{0} \Vert_{P_0}^2 \right]} =
  \bigO(n^{-\alpha}) \), for some \( \alpha\leq 1 \).
  \begin{enumerate}[label=(\alph*)]
  \item If $\alpha=1$ then
    \( \E_{P_0}{\left[ \Vert \hat{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2
      \right]} = \bigO(\log(n)n^{-1}) \).
  \item If $\alpha<1$ then
    \( \E_{P_0}{\left[ \Vert \hat{\phi}_n(\data_n^{-k}) - F_0 \Vert_{P_0}^2 \right]} =
    \bigO(n^{-\alpha}) \).
  \end{enumerate}
\end{corollary}
\begin{proof}
  See Appendix~\ref{sec:proof-proposition}.
\end{proof}

@@

Using equation~(\ref{eq:surv-def}) we may write
\( \theta_{{\tau}}(Q) = \tilde{\Psi}_{\tau}^0(\Lambda_1, \Lambda_2, \mu) \), where
\begin{equation}
  \label{eq:1}    
  \begin{split}
    \tilde{\Psi}_{\tau}^0(\Lambda_1, \Lambda_2, \mu)
    & =
      \int_{\mathcal{W}} 
      \int_0^{\tau}
      \exp{\{{-\Lambda_1(s- \mid w, 1)-\Lambda_2(s- \mid w, 1)}\}}  \Lambda_1(\diff s \mid
      w, 1)
      \mu(\diff w)
    \\
    &  \quad
      -\int_{\mathcal{W}} 
      \int_0^{\tau}
      \exp{\{{-\Lambda_1(s- \mid w, 0)-\Lambda_2(s- \mid w, 0)}\}}  \Lambda_1(\diff s \mid w, 0)
      \mu(\diff w).
  \end{split}
\end{equation}
The efficient influence function for the parameter $\tilde{\Psi}_{\tau}$ depends
on the set \( (\Lambda_1, \Lambda_2, \Gamma, \pi) \) of nuisance parameters.
We define
\begin{align*}
  \omega_a(A,W; \pi)
  &=  \frac{% (-1)^{a+1}
    \1{\{A=a\}}}{\pi(W)^{a}(1-\pi(W))^{1-a}},
  \\
  g(t, A, W; \Lambda_1, \Lambda_2)
  & = \int_0^{t}
    \exp{\{ {-\Lambda_1(s- \mid W, A)-\Lambda_2(s- \mid W, A)} \}}  \Lambda_1(\diff s \mid
    W, A),
  \\  
  M_j(\diff t \mid A, W;  \Lambda_j  )
  & = N_j(\diff t) -
    \1{\{\tilde{T} \geq t\}} \Lambda_j(\diff t \mid W, A),
    \quad j \in \{1,2\},
    \intertext{and}
    M(\diff t \mid A, W;  \Lambda_1, \Lambda_2  )
  & = M_1(\diff t \mid A, W;  \Lambda_1  ) +
    M_2(\diff t \mid A, W;  \Lambda_2  ).
\end{align*}
The efficient influence function can now be written as
\citep{van2003unified,jewell2007non,rytgaard2022targeted},
\begin{equation*}
  \psi_{\tau}(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
  = \psi_{\tau}^1(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
  - \psi_{\tau}^0(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
  -\tilde{\Psi}_{\tau}^0(\Lambda_1, \Lambda_2, \mu),
\end{equation*}
where
\begin{equation}
  \label{eq:5}
  \begin{split}
    & \psi_{\tau}^a(O; \Lambda_1, \Lambda_2, \Gamma, \pi)
    \\
    & =
      \omega_a(A,W; \pi)
      \int_0^{\tau} \exp{\{{\Gamma(t- \mid A, W)}\}}   
      M_1(\diff t \mid A, W; \Lambda_1)
    \\
    & \quad
      -
      \omega_a(A,W; \pi)
      g(\tau, A, W; \Lambda_1, \Lambda_2)
    \\
    & \quad \quad \times
      \int_0^{\tau}
      \exp{\{{[\Gamma+\Lambda_1 + \Lambda_2](t- \mid A, W)}\}}
      M(\diff t \mid A, W; \Lambda_1, \Lambda_2)
    \\
    & \quad
      +
      \omega_a(A,W; \pi)      
      \int_0^{\tau}
      g(t, A, W; \Lambda_1, \Lambda_2)
    \\
    & \quad \quad \times
      \exp{\{{[\Gamma+\Lambda_1 + \Lambda_2](t- \mid A, W)}\}}
      M(\diff t \mid A, W; \Lambda_1, \Lambda_2)
    \\
    & \quad + g(\tau, a, W; \Lambda_1, \Lambda_2).
  \end{split}
\end{equation}
Equations~(\ref{eq:1}) and~(\ref{eq:5}) allow us to construct a one-step
estimator by using the definition given in equation~(\ref{eq:one-step-def}),
which gives the estimator
\begin{equation}
  \label{eq:one-step-comp-ate}
  \begin{split}
    \hat{\Psi}_{\tau,\text{OS}}(\data_n)
    = &
        \tilde{\Psi}_{\tau}^0(\hat{a}_1(\data_n), \hat{a}_2(\data_n),
        \hat{\mu}_n(\data_n))
    \\
      &
        +
        \empmeas{[\psi_{\tau}(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
        \hat{b}_n(\data_n), \hat{\pi}_n(\data_n))]}
    \\
    = &
        \empmeas{[\psi_{\tau}^1(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
        \hat{b}_n(\data_n), \hat{\pi}_n(\data_n))]}
    \\
      &
        - \empmeas{[\psi_{\tau}^0(\blank; \hat{a}_{1,n}(\data_n), \hat{a}_{2,n}(\data_n),
        \hat{b}_n(\data_n), \hat{\pi}_n(\data_n))]}.
  \end{split}
\end{equation}

@@

\appendix

@@

\bibliography{bib.bib}
